---
title: "Homework3"
author: "Gifty Osei"
date: "2024-09-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 3c

```{r}
# Parameters
n <- 1000

# Case 1: Small p
p1 <- 0.001
lambda1 <- n * p1  # lambda = 1
k1 <- floor(lambda1 + 0.5 * sqrt(lambda1))  # k = 1

# Case 2: Larger p
p2 <- 0.01
lambda2 <- n * p2  # lambda = 10
k2 <- floor(lambda2 + 0.5 * sqrt(lambda2))  # k = 11



# Binomial CDFs
cdf_binom_case1 <- pbinom(k1, size = n, prob = p1)
cdf_binom_case2 <- pbinom(k2, size = n, prob = p2)

# Poisson CDFs
cdf_poisson_case1 <- ppois(k1, lambda = lambda1)
cdf_poisson_case2 <- ppois(k2, lambda = lambda2)



```

\begin{center} For Case 1:Small $p (p = 0.001, \lambda = 1)$, Binomial CDF at $ k=1$ is `r cdf_binom_case1 ` and Poisson CDF at $k=1$ is `r cdf_poisson_case1`.

For Case 2: Larger $p (p = 0.01, \lambda = 10)$, Binomial CDF at $k=2$ is `r cdf_binom_case2 ` and Poisson CDF at $k=2$ is `r cdf_poisson_case2`.

\end{center}


```{r, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE,include=FALSE}
# Define range for k
k_values_case1 <- 0:5
k_values_case2 <- 0:20

# Calculate CDFs for range of k
cdf_binom_range_case1 <- pbinom(k_values_case1, size = n, prob = p1)
cdf_poisson_range_case1 <- ppois(k_values_case1, lambda = lambda1)

cdf_binom_range_case2 <- pbinom(k_values_case2, size = n, prob = p2)
cdf_poisson_range_case2 <- ppois(k_values_case2, lambda = lambda2)

# Plot for Case 1
plot(k_values_case1, cdf_binom_range_case1, type = "b", pch = 19, col = "blue",
     ylim = c(0,1), xlab = "k", ylab = "CDF",
     main = "CDF Comparison: Binomial vs Poisson (p=0.001, lambda=1)")
lines(k_values_case1, cdf_poisson_range_case1, type = "b", pch = 17, col = "red")
legend("bottomright", legend = c("Binomial CDF", "Poisson CDF"),
       col = c("blue", "red"), pch = c(19,17), lty = 1)

# Plot for Case 2
plot(k_values_case2, cdf_binom_range_case2, type = "b", pch = 19, col = "blue",
     ylim = c(0,1), xlab = "k", ylab = "CDF",
     main = "CDF Comparison: Binomial vs Poisson (p=0.01, lambda=10)")
lines(k_values_case2, cdf_poisson_range_case2, type = "b", pch = 17, col = "red")
legend("bottomright", legend = c("Binomial CDF", "Poisson CDF"),
       col = c("blue", "red"), pch = c(19,17), lty = 1)

```


## Plot to show Difference


```{r plot, warning=FALSE, message=FALSE}

library(ggplot2)

# Set parameters for Binomial and Poisson distributions
n <- 1000
lambda_values <- c(5, 50)  # Small and large values of lambda
k_values <- c(floor(5 + 0.5 * sqrt(5)), floor(50 + 0.5 * sqrt(50)))  # Values for k

# Create a data frame to store CDF values
results <- data.frame(k = integer(), CDF = numeric(), Distribution = character(), Lambda = numeric())

# Calculate CDF for Binomial and Poisson distributions
for (lambda in lambda_values) {
  p <- lambda / n  # Probability of success for Binomial
  k_range <- 0:50  # Range of k values to calculate CDF

  # Calculate Binomial CDF
  binom_cdf <- pbinom(k_range, n, p)
  results <- rbind(results, data.frame(k = k_range, CDF = binom_cdf, Distribution = "Binomial", Lambda = lambda))
  
  # Calculate Poisson CDF
  poisson_cdf <- ppois(k_range, lambda)
  results <- rbind(results, data.frame(k = k_range, CDF = poisson_cdf, Distribution = "Poisson", Lambda = lambda))
}

# Plotting the CDF comparison
ggplot(results, aes(x = k, y = CDF, color = Distribution)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ Lambda, scales = "free_y", ncol = 2, labeller = labeller(Lambda = function(x) paste("λ =", x))) +
  labs(title = "CDF Comparison of Binomial and Poisson Approximations",
       x = "k", y = "CDF") +
  theme_bw() +
  theme(legend.position = "top")

```

The Poisson approximation to the Binomial distribution is more accurate when 
$\lambda$ (or $p$) is small, as illustrated in the left plot. When $\lambda$ is large, the Poisson approximation is still useful, but its accuracy diminishes as $p$ increases or $\lambda$ increases.
Therefore, The distribution $Y_n$ converges to the distribution of $Y \sim $ Poisson($\lambda$).